---
title: "Creating the Best Christmas Playlists"
output: html_notebook
---

This program finds the most popular, non-traditional versions of Christmas songs.

See the most popular songs by categories including:

- Genre
- Instrument only
- Danceability
- Acousticness
- Tempo

This can be used to create a mixed playlist or individual playlists by category.

### Install the spotifyr and curlconverter packages
```{r}
#devtools::install_github("charlie86/spotifyr")
#devtools::install_github("hrbrmstr/curlconverter")
```

```{r}
library(spotifyr)
library(curlconverter)
library(plyr)
library(dplyr)
library(jsonlite)
```

### Set Spotify credentials

```{r}
#Sys.setenv(SPOTIFY_CLIENT_ID = "your client id")
#Sys.setenv(SPOTIFY_CLIENT_SECRET = "your client secret")
```

### Search for playlists that contain a term/phrase
The `searchPlaylists` function performs a GET request that searches for Spotify playlists for a given query.
It returns a data frame with all playlist results.
```{r}
searchPlaylists <- function(query) {
  # format query
  queryUrl <- paste("'https://api.spotify.com/v1/search?q=%22", URLencode(query), "%22&type=playlist'", 
                    sep="")
  # format authorization
  authToken <- paste("'Authorization: Bearer ", token, "'", sep="")
  
  # format and make request
  req <- paste("curl -X GET ", 
               queryUrl, 
               " -H 'Accept: application/json' -H 'Content-Type: application/json' -H ",
               authToken, 
               sep = "")
  resp <- make_req(straighten(req))

  # Convert result to json
  respJson <- toJSON(content(resp[[1]](), as="parsed"), 
                     auto_unbox = TRUE, 
                     pretty=TRUE) 
  # Convert json to data frame
  respDf <- fromJSON(respJson)

  # Extract the playlist info we need
  result <- data.frame(playlist_name = respDf$playlists$items$name,
                       playlist_tracks_url = respDf$playlists$items$tracks$href,
                       playlist_num_tracks = respDf$playlists$items$tracks$total,
                       playlist_img = NA, # needed for later
                       playlist_id = respDf$playlists$items$id,
                       external_url = respDf$playlists$items$external_urls$spotify
                       )
  return(result)
}
```

```{r}
token <- get_spotify_access_token()
instrumPL <- searchPlaylists("Christmas instrumental") # Instrumental query
```


### Since this returns many playlists, get the number of followers for each playlist and only keep the playlists with the top 5 follower counts.
```{r}
# Loop through playlist_id and make GET requests to get each playlist's number of followers
# Store follower count in result$followers
result$followers <- NA
i <- 1
for (id in result$playlist_id) {
  
  followReq <- paste("curl -ik -X GET 'https://api.spotify.com/v1/playlists/", id, 
                     "' -H 'Authorization: Bearer ", token, "'", sep="")
  followResp <- make_req(straighten(followReq))
  
  # Convert result to json
  followJson <- toJSON(content(followResp[[1]](), as="parsed"), auto_unbox = TRUE, pretty=TRUE)
  # Convert json to data frame
  followDf <- fromJSON(followJson)

  # Set corresponding value in followers column
  result$followers[i] <- followDf$followers$total
  i <- i + 1
}

# Keep the top 5 playlists
result <- result %>% top_n(5, followers) %>% arrange(desc(followers))
head(result)
```



### Load in the playlist data and get track popularity

```{r}
playlists <- data.frame(playlist_name = "Christmas Instrumentals", 
                 playlist_tracks_url = "https://api.spotify.com/v1/playlists/3UwpYtXlBzw9FN24TAYksU/tracks",
                 playlist_num_tracks = 256,
                 playlist_img = NA)
# Extract the tracks
tracks <-  get_playlist_tracks(playlists) 

#NEED TO USE THIS:
#get_track_audio_features(tracks[1,])

# Get popularity
trackPop <- get_track_popularity(tracks) %>% arrange(desc(track_popularity)) 

# Join popularity df w/ track df, and select cols we want
fullTrackDf <- join(trackPop, tracks, type = "full", by = "track_uri")
fullTrackDf <- fullTrackDf %>% select(track_name, track_popularity, track_uri, playlist_name, artist_name, track_open_spotify_url)

head(fullTrackDf)
```








