---
title: "Creating the Best Christmas Playlists"
output: html_notebook
---

This program creates an ideal Christmas playlist by merging the best songs according to your interests.

Steps:  
1. Get all playlists returned by a list of custom queries (e.g. "christmas instrumentals", "christmas a cappella")  
2. Narrow the playlists to those with the top n followers  
3. Obtain the popularity of every track  
4. Perform a series of clean up steps (e.g. remove tracks below a popularity threshold, limit the number of tracks per artist)  
5. Extract audio track features  
6. Create a custom playlist depending on your desired audio features and playlist size  



### Install the spotifyr and curlconverter packages
```{r}
#devtools::install_github("charlie86/spotifyr")
#devtools::install_github("hrbrmstr/curlconverter")
```

```{r}
library(spotifyr)
library(curlconverter)
library(plyr)
library(dplyr)
library(jsonlite)
```

### Set Spotify credentials

```{r}
#Sys.setenv(SPOTIFY_CLIENT_ID = "your client id")
#Sys.setenv(SPOTIFY_CLIENT_SECRET = "your client secret")
```

### Search for playlists that contain a term/phrase
The `searchPlaylists` function performs a GET request that searches for Spotify playlists for a given query.

It returns a data frame with all playlist results.
```{r}
searchPlaylists <- function(query, token) {
  # format query
  queryUrl <- paste("'https://api.spotify.com/v1/search?q=%22", URLencode(query), "%22&type=playlist'", 
                    sep="")
  # format authorization
  authToken <- paste("'Authorization: Bearer ", token, "'", sep="")
  
  # format and make request
  req <- paste("curl -X GET ", 
               queryUrl, 
               " -H 'Accept: application/json' -H 'Content-Type: application/json' -H ",
               authToken, 
               sep = "")
  resp <- make_req(straighten(req))

  # Convert result to json
  respJson <- toJSON(content(resp[[1]](), as="parsed"), auto_unbox = TRUE, pretty=TRUE) 
  # Convert json to data frame
  respDf <- fromJSON(respJson)

  # Extract the playlist info we need
  result <- data.frame(playlist_name = respDf$playlists$items$name,
                       playlist_tracks_url = respDf$playlists$items$tracks$href,
                       playlist_num_tracks = respDf$playlists$items$tracks$total,
                       playlist_img = NA, # needed for later
                       playlist_id = respDf$playlists$items$id,
                       external_url = respDf$playlists$items$external_urls$spotify
                       )
  return(result)
}
```


### Since `searchPlaylists` returns many playlists, get the number of followers for each playlist and only keep the playlists with the top n follower counts.
```{r}
# Example: Get top 3 playlists
# getTopPLs(playlistDf, n = 3, get_spotify_access_token())
getTopPLs <- function(df, n = 5, token) {
  
  df$followers <- NA
  i <- 1
  
  # Loop through playlist_id, get each playlist's number of followers, and store result in df
  for (id in df$playlist_id) {
    followReq <- paste("curl -X GET 'https://api.spotify.com/v1/playlists/", id, 
                       "' -H 'Authorization: Bearer ", token, "'", sep="")
    followResp <- make_req(straighten(followReq))
  
    # Convert result to json
    followJson <- toJSON(content(followResp[[1]](), as="parsed"), auto_unbox = TRUE, pretty=TRUE)
    # Convert json to data frame
    followDf <- fromJSON(followJson)

    # Set corresponding value in followers column
    df$followers[i] <- followDf$followers$total
    i <- i + 1
  }
  
  return(df %>% 
           top_n(n, followers) %>% 
           arrange(desc(followers)))
}
```


### Get track popularity

```{r}
getPop <- function(topPLs) {
  # Extract the unique tracks
  tracks <-  get_playlist_tracks(topPLs) %>% 
    distinct(track_name, artist_name, .keep_all = TRUE)
  
  # Get popularity
  trackPop <- get_track_popularity(tracks) %>% 
    arrange(desc(track_popularity)) 

  # Join popularity df w/ track df, and select cols we want
  fullTrackDf <- join(trackPop, tracks, type = "full", by = "track_uri")
  fullTrackDf <- fullTrackDf %>% 
    select(track_name, track_popularity, track_uri, playlist_name, artist_name, track_open_spotify_url)

  return(fullTrackDf)
}
```

### Clean up data frame
- Remove low popularity tracks
- Remove name + artist duplicates
- Only keep a specified number of tracks w/ the same name
- Balance the number of tracks per artist

```{r}
cleanup <- function(fullTrackDf, minPop = 30, numDups = 1, songsEach = 4) {
  # Remove low popularity tracks
  fullTrackDf <- fullTrackDf %>%
    filter(track_popularity > minPop)
  
  # Remove any same name + same artist duplicates
  fullTrackDf <-  fullTrackDf %>% 
    distinct(track_name, artist_name, .keep_all = TRUE)
  
  # If multiple tracks with the same name, keep the most popular "numDups" tracks
  # Default is to only keep 1
  fullTrackDfFilt <- fullTrackDf %>%
    group_by(track_name) %>% 
    arrange(desc(track_popularity)) %>%
    top_n(numDups, track_popularity)

  # For every artist, keep their top "songsEach" songs (or more if tied). 
  # Default is top 4 songs.
  fullTrackDfFilt <- fullTrackDfFilt %>% 
    group_by(artist_name) %>% 
    top_n(songsEach, track_popularity)

  return(fullTrackDfFilt)
}
```


### Run all functions
Returns a data frame of all tracks with their popularity info
```{r}
# Functions takes in a list of search queries and an optional list of specific playlists to include
createPlaylist <- function(queries, playlists = NULL) {
  allTracks <- data.frame()
  
  for (query in queries) {
    # Get all playlists matching search term
    allPlaylists <- searchPlaylists(query, get_spotify_access_token())
    
    # Filter by playlists with most followers
    topPlaylists <- getTopPLs(allPlaylists, n = 5, get_spotify_access_token())
    
    # Get track popularity
    trackDf <- getPop(topPlaylists)
    
    # If playlists arg is supplied, get their track popularities and join all
    if (!is.null(playlists)) {
       for (pl in playlists) {
         trackDf <- rbind(trackDf, getPop(pl))
       }
    }
    
    allTracks <- rbind(allTracks, trackDf)
  }
  
  return(allTracks)
}
```

### Test it out! Create a playlist of 100 top instrumental songs, keeping the energy audio feature >= 0.27.

```{r}
# First set queries and use my own instrumental playlist
queries <- c("Christmas instrumental", "upbeat christmas instrumental")
myInstPl <- get_user_playlists('corinneleopold') %>% filter(playlist_name == "Instrumental/Background Xmas Mix")

# Get results
results <- createPlaylist(queries, list(myInstPl))
trackDf <- cleanup(results, minPop = 30, numDups = 2) # Cleanup results with adjustable parameters
trackDf <- as.data.frame(trackDf)

# Merge results w/ audio features
trackFeats <- get_track_audio_features(trackDf, get_spotify_access_token())
trackDf <- join(trackDf, trackFeats, by = "track_uri")
```

### To get my final playlist, I want to sample from the track dataframe until I get a playlist w/ average energy >= 0.27 and average popularity >= 40.

```{r}
getFinalPlaylist <- function(df) {
  aveEnergy <- 0
  avePop <- 0
  
  while (aveEnergy < 0.27 | avePop < 40) {
    songs <- sample_n(df, 100, weight = df$wt)
    aveEnergy <- mean(songs$energy)
    avePop <- mean(songs$track_popularity)
  }
  return(songs)
}

# Add probability weights to dataframe so that songs w/ higher energy have a higher prob
trackDf <- trackDf %>%
  mutate(wt = ifelse(energy > 0.23, .75, .25))
```

```{r}
# Get final playlist
finalPlaylist <- getFinalPlaylist(trackDf)
head(arrange(finalPlaylist, desc(track_popularity)), 20)
```

### Add tracks to Spotify playlist
1. In Spotify, create an empty playlist
2. Open web browser and go to the following url in order to allow the scope "playlist-modify-public":
https://accounts.spotify.com/en/authorize?client_id=<your id>&redirect_uri=<your app uri>
&scope=playlist-modify-public%20playlist-modify-public&response_type=token&state=123  


    The app's redirect_uri can be set to "http://localhost:8888/callback" in the Spotify developer dashboard.  

3. Step 2 will return an access token in the redirect url that you will copy and paste here to use in the "token" argument.
```{r}

# PASTE TOKEN HERE
token <- ""

# Makes request to add songs, specified as a comma separated list in trackUris
makeReq <- function(trackUris, username, playlist, token) {
  
  authToken <- paste("'Authorization: Bearer ", token, "'", sep="")
  url <- paste("'https://api.spotify.com/v1/users/", username, "/playlists/", 
               playlist$playlist_uri, "/tracks?uris=", trackUris, "'", sep="")
  
  req <- paste("curl -i -X POST ",
                 url,
                 " -H ", authToken, 
                 " -H 'Content-Type: application/json'",
                 " -H 'Accept: application/json'",
                 sep = "")
    resp <- make_req(straighten(req))
  
    respJson <- toJSON(content(resp[[1]](), as="parsed"), auto_unbox = TRUE, pretty=TRUE) 
    return(respJson)
  }

makePlaylist <- function(username, spotifyPlaylistName, finalPlaylist, token) {
  
  # Paste "spotify:track:" in front of all track uris
  finalPlaylist$track_uri <- paste0("spotify:track:", finalPlaylist$track_uri)
  
  # Split df because we can only add a number of songs at a time
  # Make track uris comma separated
  playlist1 <- finalPlaylist %>% slice(1:10)
  trackUris1 <- paste0(playlist1$track_uri, collapse = ",")
  playlist2 <- finalPlaylist %>% slice(11:50)
  trackUris2 <- paste0(playlist2$track_uri, collapse = ",")
  playlist3 <- finalPlaylist %>% slice(51:100)
  trackUris3 <- paste0(playlist3$track_uri, collapse = ",")

  # Get the playlist you wish to add these songs to (or create a new empty one)
  playlist <- get_user_playlists(username) %>% filter(playlist_name == spotifyPlaylistName)
  
  # Add songs to playlist
  makeReq(trackUris1, username, playlist, token)
  makeReq(trackUris2, username, playlist, token)
  makeReq(trackUris3, username, playlist, token)
}
```

```{r}
makePlaylist(username = "corinneleopold", 
             spotifyPlaylistName = "New Xmas Playlist",
             finalPlaylist, 
             token)
```

